### 2024-04-29 thesis/input_files/analysis.tex
correct names in contents,list of fig and tables

\subsection{Tbats}
Next, we tried with the autoregressive approach. Unfortunately, AR, ARIMA and SARIMA models did not perform as expected, their output was slightly better than the one of linear regression. This is probably due to the fact that, the data considered entails two kinds of seasonalities, while ARIMA models can only handle one at a time. We remind the reader, that the electricity load time series involves both daily and weekly seasonalities. Hence, the need for a more advanced time series model. The Tbats \cite{de2011forecasting} model is a forecasting method capable of handling complex patterns in the data. Its name stands for trigonometric seasonality, Box-Cox transformation, ARMA errors, trend and seasonal components. 
Tbats forecast is visualized in figure \ref{fig:tbats_price}, meanwhile its RMSE is 15.08.
\begin{figure}[!h]
    \includegraphics[width=\textwidth]{images/tbats_price.png}
    \caption{Tbats prediction}
    \label{fig:tbats_price}
\end{figure}
From the plot we see that Tbats is capable of catching the lows and average trend, conversely it has some difficulties handling the price peaks.
\subsection{Prophet}
Following, the prophet model has been considered. 
We get started by considering the base implementation. In such setting, prophet takes in as input the time series object and learns its data generating process.
This method achieves a RMSE of 23.96, its prediction is visualized in figure \ref{fig:prophet_price_1}.
\begin{figure}[!h]
    \includegraphics[width=\textwidth]{images/prophet_price_1.png}
    \caption{Prophet prediction}
    \label{fig:prophet_price_1}
\end{figure}
What can be seen is that, prophet models correctly the average trend but does not model peaks and lows precisely. 
Next, a more complex model was trained. We added the weather temperature, the square of it and the categorical variable for holidays effect as regressors. Furthermore, we also applied a log transformation to the dependet variable. Doing so, RMSE went down to 10.29. Forecast is visualized in \ref{fig:prophet_price2}, moreover, figure \ref{fig:prophet_price2.1} and figure \ref{fig:prophet_price2.2} break down the trend and the seasonalities respectively.
\begin{figure}[!h]
    \includegraphics[width=\textwidth]{images/prophet_price2.png}
    \caption{Prophet v2 prediction}
    \label{fig:prophet_price2}
\end{figure}

@@

\subsection{LSTM}
We used the torch library in order to code our LSTM regressor.
Various hyperparameters combinations were tried during hyperparameter tuning. The final hyperparameters we set on follows.
\begin{itemize}
    \item hidden\_size= 64
    \item num\_layers= 2
    \item output\_size= 1
    \item num\_epochs= 30
    \item learning\_rate= 0.001
    \item batch\_size= 32
    \item window\_size= 24
\end{itemize}
LSTM forecast is reported in table \ref{fig:lstm_price}, it achieves a RMSE of 9.1782
\begin{figure}[!h]
    \includegraphics[width=\textwidth]{images/lstm_price.png}
    \caption{LSTM prediction}
    \label{fig:lstm_price}
\end{figure}

@@

\begin{table}[!ht]
    \caption{RMSE scores}
    \label{tab:point_RMSE}
    \begin{adjustbox}{width=\textwidth, height=3cm}
        \begin{tabular}{lrrrrrrrrrrrrrrr}
            \toprule
            Method/RMSE & Task 1 & Task 2 & Task 3 & Task 4 & Task 5 & Task 6 & Task 7 & Task 8 & Task 9 & Task 10 & Task 11 & Task 12 & Task 13 & Task 14 & Task 15 \\
            \midrule
            MLR & 30.5892 & 29.0347 & 75.1682 & 63.1376 & 42.1156 & 36.6233 & 39.2496 & 38.8286 & 47.1307 & 68.0494 & 61.6005 & 30.3751 & 34.5523 & 33.4287 & 33.5581 \\
            TBATS & 15.0886 & 31.9720 & 99.1260 & 84.4890 & 51.9740 & 34.9055 & 18.4445 & 38.3246 & 74.1117 & 98.6600 & 84.3050 & 38.8433 & 16.6080 & 28.9902 & 41.6194 \\
            Prophet & 10.2936 & 14.4358 & 38.7551 & 63.4787 & 19.7474 & 17.5065 & 12.6926 & 14.2665 & 17.5466 & 23.5944 & 43.6666 & 20.8637 & 16.9493 & 19.1626 & 23.3889 \\
            KNN & 12.5429 & 11.6699 & 24.5057 & 18.3310 & 13.1821 & \textbf{11.9238} & 12.0044 & 14.5165 & 16.3132 & 15.1831 & 37.6457 & 16.4690 & 12.0324 & 11.3102 & 14.1717 \\
            SVR & 23.2421 & 25.8525 & 78.8043 & 67.6209 & 42.3748 & 32.7845 & 30.5971 & 35.3660 & 55.4213 & 77.9660 & 68.5799 & 30.2451 & 27.0345 & 28.3652 & 32.1480 \\
            LSTM & 9.1782 & 12.0669 & \textbf{22.7048} & \textbf{16.2087} & 16.0964 & 12.7936 & \textbf{10.8559} & 14.6173 & 19.7303 & 18.0200 & 43.2051 & 17.1856 & 10.3106 & 12.1347 & 17.5849 \\
            KrnRidge & 9.8619 & 11.6101 & 37.7824 & 35.2459 & \textbf{12.5160} & 14.5911 & 12.8791 & 17.4385 & 16.1131 & 17.1938 & 37.6961 & 14.0076 & 9.8441 & 10.7491 & 13.2975 \\
            KrnSVR & \textbf{9.1028} & \textbf{11.3117} & 22.9281 & 19.2132 & 12.6331 & 12.6018 & 11.3537 & \textbf{12.9506} & \textbf{14.9731} & \textbf{11.7765} & \textbf{37.3797} & \textbf{13.2694} & 
            \textbf{8.8522} & \textbf{10.6185} & \textbf{13.2602} \\
            \bottomrule
            \end{tabular}            
    \end{adjustbox}
\end{table}

The MAE scores are contained in table \ref{tab:point_MAE}. Similarly to above, we can conclude that kernel methods stand out for achieving also the lowest MAE score.
\begin{table}[!ht]
    \caption{MAE scores}
    \label{tab:point_MAE}
    \begin{adjustbox}{width=\textwidth, height=3cm}
        \begin{tabular}{lrrrrrrrrrrrrrrr}
            \toprule
             Method/MAE & Task 1 & Task 2 & Task 3 & Task 4 & Task 5 & Task 6 & Task 7 & Task 8 & Task 9 & Task 10 & Task 11 & Task 12 & Task 13 & Task 14 & Task 15 \\
            \midrule
            MLR & 27.7219 & 24.3823 & 65.4016 & 52.3873 & 34.8465 & 31.2663 & 35.9420 & 34.7316 & 37.0273 & 54.7075 & 48.9327 & 26.1975 & 31.8720 & 29.2306 & 28.3428 \\
            TBATS & 10.5408 & 23.8417 & 91.3153 & 74.5734 & 40.3258 & 26.0092 & 14.2141 & 24.7818 & 62.0452 & 86.7125 & 72.6912 & 28.4235 & 11.1184 & 21.1038 & 31.8555 \\
            Prophet & 8.6139 & 11.4307 & 28.3509 & 46.4480 & 15.6290 & 14.1278 & 10.1574 & 11.2360 & 14.0072 & 18.9565 & 27.1142 & 16.9678 & 14.6615 & 16.1750 & 18.3043 \\
            KNN & 9.4003 & 9.1126 & 19.6213 & 14.4906 & 10.6461 & \textbf{9.4354} & 8.7441 & 10.0065 & 12.4714 & 11.2006 & 20.3087 & 12.5416 & 9.2558 & 8.5937 & 10.5451 \\
            SVR & 20.2443 & 21.2526 & 69.3704 & 57.1107 & 34.3651 & 27.5875 & 27.4651 & 29.4204 & 43.7009 & 64.3522 & 55.4307 & 25.1080 & 24.0294 & 24.4049 & 26.5869 \\
            LSTM & 7.5217 & 9.4767 & \textbf{17.9851} & \textbf{13.2158} & 11.8118 & 9.8607 & \textbf{8.0561} & 10.8480 & 15.1008 & 13.8108 & 25.9906 & 13.3588 & 8.1526 & 9.8000 & 14.3186 \\
            KrnRidge & 7.5965 & 9.2130 & 28.2484 & 24.2366 & \textbf{10.0589} & 10.7758 & 9.5788 & 11.2166 & 12.3427 & 12.3306 & 20.0213 & 10.8434 & 7.1625 & \textbf{7.9710} & 10.0767 \\
            KrnSVR & \textbf{6.9581} & \textbf{8.6989} & 18.7739 & 15.1398 & 10.1395 & 9.5927 & 8.5731 & \textbf{9.0033} & \textbf{11.0618} & \textbf{8.5486} & \textbf{19.7383} & \textbf{10.5545} & \textbf{7.0395} & 8.1926 & \textbf{9.8125} \\
            \bottomrule
            \end{tabular}            
    \end{adjustbox}            
\end{table}

Finally table \ref{tab:point_MAPE} reports the MAPE score for each method. What can be concluded is that kernel regressors are also the best in terms of MAPE score; with KrnSVR being the top one.
\begin{table}[!ht]
    \caption{MAPE scores}
    \label{tab:point_MAPE}
    \begin{adjustbox}{width=\textwidth, height=3cm}
        \begin{tabular}{lrrrrrrrrrrrrrrr}
            \toprule
             Method/MAPE & Task 1 & Task 2 & Task 3 & Task 4 & Task 5 & Task 6 & Task 7 & Task 8 & Task 9 & Task 10 & Task 11 & Task 12 & Task 13 & Task 14 & Task 15 \\
            \midrule
            MLR & 0.2470 & 0.1935 & 0.3030 & 0.2601 & 0.2474 & 0.2667 & 0.3498 & 0.3058 & 0.1954 & 0.2432 & 0.4234 & 0.2114 & 0.2926 & 0.2502 & 0.2145 \\
            TBATS & 0.0833 & 0.1650 & 0.4285 & 0.3731 & 0.2482 & 0.1849 & 0.1234 & 0.1634 & 0.3181 & 0.4021 & 0.4880 & 0.1761 & 0.0905 & 0.1555 & 0.2058 \\
            Prophet & 0.0737 & 0.0854 & 0.1324 & 0.2453 & 0.1134 & 0.1169 & 0.0938 & 0.0912 & 0.0880 & 0.0992 & 0.3757 & 0.1389 & 0.1343 & 0.1356 & 0.1370 \\
            KNN & 0.0778 & 0.0682 & 0.0924 & 0.0761 & 0.0751 & 0.0740 & 0.0784 & 0.0744 & 0.0725 & 0.0555 & \textbf{0.3115} & 0.0888 & 0.0816 & 0.0699 & 0.0765 \\
            SVR & 0.1808 & 0.1640 & 0.3211 & 0.2829 & 0.2317 & 0.2260 & 0.2676 & 0.2451 & 0.2215 & 0.2867 & 0.4354 & 0.1870 & 0.2216 & 0.2044 & 0.1933 \\
            LSTM & 0.0642 & 0.0706 & 0.0896 & \textbf{0.0722} & 0.0798 & 0.0772 & \textbf{0.0710} & 0.0854 & 0.0909 & 0.0741 & 0.3578 & 0.0976 & 0.0719 & 0.0778 & 0.1037 \\
            KrnRidge & 0.0625 & 0.0685 & 0.1281 & 0.1207 & 0.0718 & 0.0797 & 0.0812 & 0.0760 & 0.0696 & 0.0587 & 0.3127 & 0.0779 & \textbf{0.0591} & \textbf{0.0637} & 0.0722 \\
            KrnSVR & \textbf{0.0568} & \textbf{0.0642} & \textbf{0.0892} & 0.0788 & \textbf{0.0714} & \textbf{0.0724} & 0.0740 & \textbf{0.0665} & \textbf{0.0629} & \textbf{0.0428} & 0.3122 & \textbf{0.0781} & 0.0607 & 0.0646 & \textbf{0.0713} \\
            \bottomrule
            \end{tabular}            
    \end{adjustbox}
\end{table}

>>>>

\subsection{Trigonometric seasonality Box-Cox transformation ARMA errors trend and seasonal components}
Next, we tried with the autoregressive approach. Unfortunately, AR, ARIMA and SARIMA models did not perform as expected, their output was slightly better than the one of linear regression. This is probably due to the fact that, the data considered entails two kinds of seasonalities, while ARIMA models can only handle one at a time. We remind the reader, that the electricity load time series involves both daily and weekly seasonalities. Hence, the need for a more advanced time series model. The Tbats \cite{de2011forecasting} model is a forecasting method capable of handling complex patterns in the data. Its name stands for trigonometric seasonality, Box-Cox transformation, ARMA errors, trend and seasonal components. 
Tbats forecast is visualized in figure \ref{fig:tbats_price}, meanwhile its RMSE is 15.08.
\begin{figure}[!h]
    \includegraphics[width=\textwidth]{images/tbats_price.png}
    \caption{Tbats prediction}
    \label{fig:tbats_price}
\end{figure}
From the plot we see that Tbats is capable of catching the lows and average trend, conversely it has some difficulties handling the price peaks.
\subsection{Prophet}
Following, the prophet model has been considered. 
We get started by considering the base implementation. In such setting, prophet takes in as input the time series object and learns its data generating process.
This method achieves a RMSE of 23.96, its prediction is visualized in figure \ref{fig:prophet_price_1}.
\begin{figure}[!h]
    \includegraphics[width=\textwidth]{images/prophet_price_1.png}
    \caption{Prophet prediction}
    \label{fig:prophet_price_1}
\end{figure}
What can be seen is that, prophet models correctly the average trend but does not model peaks and lows precisely. 
Next, a more complex model was trained. We added the weather temperature, the square of it and the categorical variable for holidays effect as regressors. Furthermore, we also applied a log transformation to the dependet variable. Doing so, RMSE went down to 10.29. Forecast is visualized in \ref{fig:prophet_price2}, moreover, figure \ref{fig:prophet_price2.1} and figure \ref{fig:prophet_price2.2} break down the trend and the seasonalities respectively.
\begin{figure}[!h]
    \includegraphics[width=\textwidth]{images/prophet_price2.png}
    \caption{Prophet v2 prediction}
    \label{fig:prophet_price2}
\end{figure}

@@

\subsection{Long short term memory}
We used the torch library in order to code our LSTM regressor.
Various hyperparameters combinations were tried during hyperparameter tuning. The final hyperparameters we set on follows.
\begin{itemize}
    \item hidden\_size= 64
    \item num\_layers= 2
    \item output\_size= 1
    \item num\_epochs= 30
    \item learning\_rate= 0.001
    \item batch\_size= 32
    \item window\_size= 24
\end{itemize}
LSTM forecast is reported in table \ref{fig:lstm_price}, it achieves a RMSE of 9.1782
\begin{figure}[!h]
    \includegraphics[width=\textwidth]{images/lstm_price.png}
    \caption{Long short term memory prediction}
    \label{fig:lstm_price}
\end{figure}

@@

\begin{table}[!ht]
    \caption{Root mean squared errors}
    \label{tab:point_RMSE}
    \begin{adjustbox}{width=\textwidth, height=3cm}
        \begin{tabular}{lrrrrrrrrrrrrrrr}
            \toprule
            Method/RMSE & Task 1 & Task 2 & Task 3 & Task 4 & Task 5 & Task 6 & Task 7 & Task 8 & Task 9 & Task 10 & Task 11 & Task 12 & Task 13 & Task 14 & Task 15 \\
            \midrule
            MLR & 30.5892 & 29.0347 & 75.1682 & 63.1376 & 42.1156 & 36.6233 & 39.2496 & 38.8286 & 47.1307 & 68.0494 & 61.6005 & 30.3751 & 34.5523 & 33.4287 & 33.5581 \\
            TBATS & 15.0886 & 31.9720 & 99.1260 & 84.4890 & 51.9740 & 34.9055 & 18.4445 & 38.3246 & 74.1117 & 98.6600 & 84.3050 & 38.8433 & 16.6080 & 28.9902 & 41.6194 \\
            Prophet & 10.2936 & 14.4358 & 38.7551 & 63.4787 & 19.7474 & 17.5065 & 12.6926 & 14.2665 & 17.5466 & 23.5944 & 43.6666 & 20.8637 & 16.9493 & 19.1626 & 23.3889 \\
            KNN & 12.5429 & 11.6699 & 24.5057 & 18.3310 & 13.1821 & \textbf{11.9238} & 12.0044 & 14.5165 & 16.3132 & 15.1831 & 37.6457 & 16.4690 & 12.0324 & 11.3102 & 14.1717 \\
            SVR & 23.2421 & 25.8525 & 78.8043 & 67.6209 & 42.3748 & 32.7845 & 30.5971 & 35.3660 & 55.4213 & 77.9660 & 68.5799 & 30.2451 & 27.0345 & 28.3652 & 32.1480 \\
            LSTM & 9.1782 & 12.0669 & \textbf{22.7048} & \textbf{16.2087} & 16.0964 & 12.7936 & \textbf{10.8559} & 14.6173 & 19.7303 & 18.0200 & 43.2051 & 17.1856 & 10.3106 & 12.1347 & 17.5849 \\
            KrnRidge & 9.8619 & 11.6101 & 37.7824 & 35.2459 & \textbf{12.5160} & 14.5911 & 12.8791 & 17.4385 & 16.1131 & 17.1938 & 37.6961 & 14.0076 & 9.8441 & 10.7491 & 13.2975 \\
            KrnSVR & \textbf{9.1028} & \textbf{11.3117} & 22.9281 & 19.2132 & 12.6331 & 12.6018 & 11.3537 & \textbf{12.9506} & \textbf{14.9731} & \textbf{11.7765} & \textbf{37.3797} & \textbf{13.2694} & 
            \textbf{8.8522} & \textbf{10.6185} & \textbf{13.2602} \\
            \bottomrule
            \end{tabular}            
    \end{adjustbox}
\end{table}

The MAE scores are contained in table \ref{tab:point_MAE}. Similarly to above, we can conclude that kernel methods stand out for achieving also the lowest MAE score.
\begin{table}[!ht]
    \caption{Mean absolute errors}
    \label{tab:point_MAE}
    \begin{adjustbox}{width=\textwidth, height=3cm}
        \begin{tabular}{lrrrrrrrrrrrrrrr}
            \toprule
             Method/MAE & Task 1 & Task 2 & Task 3 & Task 4 & Task 5 & Task 6 & Task 7 & Task 8 & Task 9 & Task 10 & Task 11 & Task 12 & Task 13 & Task 14 & Task 15 \\
            \midrule
            MLR & 27.7219 & 24.3823 & 65.4016 & 52.3873 & 34.8465 & 31.2663 & 35.9420 & 34.7316 & 37.0273 & 54.7075 & 48.9327 & 26.1975 & 31.8720 & 29.2306 & 28.3428 \\
            TBATS & 10.5408 & 23.8417 & 91.3153 & 74.5734 & 40.3258 & 26.0092 & 14.2141 & 24.7818 & 62.0452 & 86.7125 & 72.6912 & 28.4235 & 11.1184 & 21.1038 & 31.8555 \\
            Prophet & 8.6139 & 11.4307 & 28.3509 & 46.4480 & 15.6290 & 14.1278 & 10.1574 & 11.2360 & 14.0072 & 18.9565 & 27.1142 & 16.9678 & 14.6615 & 16.1750 & 18.3043 \\
            KNN & 9.4003 & 9.1126 & 19.6213 & 14.4906 & 10.6461 & \textbf{9.4354} & 8.7441 & 10.0065 & 12.4714 & 11.2006 & 20.3087 & 12.5416 & 9.2558 & 8.5937 & 10.5451 \\
            SVR & 20.2443 & 21.2526 & 69.3704 & 57.1107 & 34.3651 & 27.5875 & 27.4651 & 29.4204 & 43.7009 & 64.3522 & 55.4307 & 25.1080 & 24.0294 & 24.4049 & 26.5869 \\
            LSTM & 7.5217 & 9.4767 & \textbf{17.9851} & \textbf{13.2158} & 11.8118 & 9.8607 & \textbf{8.0561} & 10.8480 & 15.1008 & 13.8108 & 25.9906 & 13.3588 & 8.1526 & 9.8000 & 14.3186 \\
            KrnRidge & 7.5965 & 9.2130 & 28.2484 & 24.2366 & \textbf{10.0589} & 10.7758 & 9.5788 & 11.2166 & 12.3427 & 12.3306 & 20.0213 & 10.8434 & 7.1625 & \textbf{7.9710} & 10.0767 \\
            KrnSVR & \textbf{6.9581} & \textbf{8.6989} & 18.7739 & 15.1398 & 10.1395 & 9.5927 & 8.5731 & \textbf{9.0033} & \textbf{11.0618} & \textbf{8.5486} & \textbf{19.7383} & \textbf{10.5545} & \textbf{7.0395} & 8.1926 & \textbf{9.8125} \\
            \bottomrule
            \end{tabular}            
    \end{adjustbox}            
\end{table}

Finally table \ref{tab:point_MAPE} reports the MAPE score for each method. What can be concluded is that kernel regressors are also the best in terms of MAPE score; with KrnSVR being the top one.
\begin{table}[!ht]
    \caption{Mean absolute percentage errors}
    \label{tab:point_MAPE}
    \begin{adjustbox}{width=\textwidth, height=3cm}
        \begin{tabular}{lrrrrrrrrrrrrrrr}
            \toprule
             Method/MAPE & Task 1 & Task 2 & Task 3 & Task 4 & Task 5 & Task 6 & Task 7 & Task 8 & Task 9 & Task 10 & Task 11 & Task 12 & Task 13 & Task 14 & Task 15 \\
            \midrule
            MLR & 0.2470 & 0.1935 & 0.3030 & 0.2601 & 0.2474 & 0.2667 & 0.3498 & 0.3058 & 0.1954 & 0.2432 & 0.4234 & 0.2114 & 0.2926 & 0.2502 & 0.2145 \\
            TBATS & 0.0833 & 0.1650 & 0.4285 & 0.3731 & 0.2482 & 0.1849 & 0.1234 & 0.1634 & 0.3181 & 0.4021 & 0.4880 & 0.1761 & 0.0905 & 0.1555 & 0.2058 \\
            Prophet & 0.0737 & 0.0854 & 0.1324 & 0.2453 & 0.1134 & 0.1169 & 0.0938 & 0.0912 & 0.0880 & 0.0992 & 0.3757 & 0.1389 & 0.1343 & 0.1356 & 0.1370 \\
            KNN & 0.0778 & 0.0682 & 0.0924 & 0.0761 & 0.0751 & 0.0740 & 0.0784 & 0.0744 & 0.0725 & 0.0555 & \textbf{0.3115} & 0.0888 & 0.0816 & 0.0699 & 0.0765 \\
            SVR & 0.1808 & 0.1640 & 0.3211 & 0.2829 & 0.2317 & 0.2260 & 0.2676 & 0.2451 & 0.2215 & 0.2867 & 0.4354 & 0.1870 & 0.2216 & 0.2044 & 0.1933 \\
            LSTM & 0.0642 & 0.0706 & 0.0896 & \textbf{0.0722} & 0.0798 & 0.0772 & \textbf{0.0710} & 0.0854 & 0.0909 & 0.0741 & 0.3578 & 0.0976 & 0.0719 & 0.0778 & 0.1037 \\
            KrnRidge & 0.0625 & 0.0685 & 0.1281 & 0.1207 & 0.0718 & 0.0797 & 0.0812 & 0.0760 & 0.0696 & 0.0587 & 0.3127 & 0.0779 & \textbf{0.0591} & \textbf{0.0637} & 0.0722 \\
            KrnSVR & \textbf{0.0568} & \textbf{0.0642} & \textbf{0.0892} & 0.0788 & \textbf{0.0714} & \textbf{0.0724} & 0.0740 & \textbf{0.0665} & \textbf{0.0629} & \textbf{0.0428} & 0.3122 & \textbf{0.0781} & 0.0607 & 0.0646 & \textbf{0.0713} \\
            \bottomrule
            \end{tabular}            
    \end{adjustbox}
\end{table}

### 2024-04-29 thesis/input_files/energymarket.tex
correct names in contents,list of fig and tables

\begin{figure}[!h]
    \includegraphics[width=\textwidth]{images/pool_vs_echange.png}
    \caption{Pool market vs power echange \cite{weron2006modeling}}
    \label{fig:pool_vs_echange}
\end{figure}

@@

\subsection{Nodal and Zonal Pricing}
Grid configuration and physical limits on electricity lines may lead to congestion. Local marginal price (LMP) and zonal market clearing pricing (ZMCP) are two types of pricing schemes utilized to handle it. The former is adopted in EU contries while the latter is used in the US, figure \ref{fig:eu_zonal} and figure \ref{fig:us_local}.
With ZMCP, prices may differ between zones but are the same within the same area.
On the other hand, LMP is made up by summing the transmission congestion cost, generation marginal cost and the cost of marginal losses at different buses; buses is where a electicity line or several lines are connected.
\begin{figure}[!h]
    \includegraphics[width=\textwidth]{images/eu_zonal.png}
    \caption{Zonal pricing EU \cite{eu_zonal}}
    \label{fig:eu_zonal}
\end{figure}

>>>>

\begin{figure}[!h]
    \includegraphics[width=\textwidth]{images/pool_vs_echange.png}
    \caption{Pool market versus power echange \cite{weron2006modeling}}
    \label{fig:pool_vs_echange}
\end{figure}

@@

\subsection{Nodal and zonal pricing}
Grid configuration and physical limits on electricity lines may lead to congestion. Local marginal price (LMP) and zonal market clearing pricing (ZMCP) are two types of pricing schemes utilized to handle it. The former is adopted in EU contries while the latter is used in the US, figure \ref{fig:eu_zonal} and figure \ref{fig:us_local}.
With ZMCP, prices may differ between zones but are the same within the same area.
On the other hand, LMP is made up by summing the transmission congestion cost, generation marginal cost and the cost of marginal losses at different buses; buses is where a electicity line or several lines are connected.
\begin{figure}[!h]
    \includegraphics[width=\textwidth]{images/eu_zonal.png}
    \caption{Zonal pricing EU \cite{eu_zonal}}
    \label{fig:eu_zonal}
\end{figure}

### 2024-04-29 thesis/input_files/implementation.tex
April 28 2024
correct names in contents,list of fig and tables

% - see whether something can be parallelized

@@

\subsection{LSTM}

@@

\section{Probabilistic forecasting}
\subsection{Linear quantile regression}
The implementation of statsmodels.regression.quantile\_regression.QuantReg has been used.
The model is fitted through iterative reweighted least squares.
\subsection{Quantile gradient boosting machine}
The implementation of sklearn.ensemble.GradientBoostingRegressor has been used.
\subsection{Quantile forest}
The implementation of quantile\_forest.RandomForestQuantileRegressor has been used. This estimator is compatible with scikit-learn.
\subsection{Kernel quantile regression}
Kernel quantile regression had no previously implemented open source library, thus the need of implementing our own version.
\\
The scikit-learn team provides a project template for the creation of estimators compatible with scikit-learn functionalities. Therefore, the KQR class is derived from the scikit-learn BaseEstimator and the mixin class RegressorMixin.
Our KQR class is initialized by providing a quantile, the regularization term C and the rbf kernel bandwidth gamma; the latter are the two hyperparameters of our custom estimator.
\\
In the fit method, we set up and solve the convex optimisation problem by using the cvxopt library. 
When using this library, it is important to keep two things in mind. First this library assumes the quadratic term of the optimisation problem to be multiplied by the 0.5 factor, thus we just have to provide the Q matrix with no 0.5 in front.
Secondly, in order to specify multiple inequalities we have to stack them an provide as a unique matrix.
\\
Once a solution to the convex problem has been found, we create a mask for the support vectors of the estimator in order to estimate the constant term of our kernel quantile regressor.
\\
In the predict method, we pass a matrix X\_eval of independent variables, next we compute the kernel matrix between X\_train and X\_eval and obtain y\_eval with the formula $y=\alpha^\intercal K+b$.
\\
This estimator is compatible with useful scikit-learn in built methods like gridsearch, crossvalidation and scoring rules.
\\
Up to now, there are only two open source implementation of the quantile kernel regression. Nevertheless they are both in R, that is there exists no python, matlab or julia open source implementation. Following are reported the results of a comparative study between our own implementation and the one of the R library kernlab
% (i guess it is in C or C++ and then binded to R).
\subsubsection{Python vs R implementation}
In this section, a comparison study has been carried out in order to inspect the competitiveness of our implementation with the existing one for the R programming language.

>>>>

@@

\subsection{LSTM}
The LSTM predictor has been built using the \href{https://pytorch.org}{torch} library.

@@

\section{Probabilistic forecasting}
\subsection{Linear quantile regression}
The implementation of statsmodels.regression.quantile\_regression.QuantReg has been used.
The model is fitted through iterative reweighted least squares.
\subsection{Quantile gradient boosting machine}
The implementation of sklearn.ensemble.GradientBoostingRegressor has been used.
\subsection{Quantile forest}
The implementation of quantile\_forest.RandomForestQuantileRegressor has been used. This estimator is compatible with scikit-learn.
\subsection{Kernel quantile regression}
Kernel quantile regression had no previously implemented open source library, thus the need of implementing our own version.
\\
The scikit-learn team provides a project template for the creation of estimators compatible with scikit-learn functionalities. Therefore, the KQR class is derived from the scikit-learn BaseEstimator and the mixin class RegressorMixin.
Our KQR class is initialized by providing a quantile, the regularization term C and the rbf kernel bandwidth gamma; the latter are the two hyperparameters of our custom estimator.
\\
In the fit method, we set up and solve the convex optimisation problem by using the cvxopt library. 
When using this library, it is important to keep two things in mind. First this library assumes the quadratic term of the optimisation problem to be multiplied by the 0.5 factor, thus we just have to provide the Q matrix with no 0.5 in front.
Secondly, in order to specify multiple inequalities we have to stack them an provide as a unique matrix.
\\
Once a solution to the convex problem has been found, we create a mask for the support vectors of the estimator in order to estimate the constant term of our kernel quantile regressor.
\\
In the predict method, we pass a matrix X\_eval of independent variables, next we compute the kernel matrix between X\_train and X\_eval and obtain y\_eval with the formula $y=\alpha^\intercal K+b$.
\\
This estimator is compatible with useful scikit-learn in built methods like gridsearch, crossvalidation and scoring rules.
\\
Up to now, there are only two open source implementation of the quantile kernel regression. Nevertheless they are both in R, that is there exists no python, matlab or julia open source implementation. Following are reported the results of a comparative study between our own implementation and the one of the R library kernlab
% (i guess it is in C or C++ and then binded to R).
\subsubsection{Python versus R implementation}
In this section, a comparison study has been carried out in order to inspect the competitiveness of our implementation with the existing one for the R programming language.
